----
Libraries
----

```{r}
library(tidyverse)
library(janitor)
library(stringr)
library(readxl)
```

----
Load datasets
----

Discharges By Health Board
```{r}
discharges_by_healthboard <- read_csv("../../data/raw_data/stroke_activitybyhbr.csv") %>% 
# Remove Qualifiers
  select(-contains("QF")) %>% 
# Clean Names
  clean_names() %>% 
# Make Readable
  rename(healthboard_area = hbr) %>% 
# Remove Unused Variable
  select(-admission_type)
```

Mortality By Health Board
```{r}
mortality_by_healthboard <- read_csv("../../data/raw_data/stroke_mortalitybyhbr.csv") %>% 
# Remove Qualifer
  select(-contains("QF")) %>% 
# Clean names
  clean_names() %>% 
# Make Readable
  rename(healthboard_area = hbr)
```

Hospital Stays By Health Board
```{r}
hospitalstays_by_healthboard <- read_xlsx("../../data/raw_data/Diagnosis-by-Health-Board-of-Residence-Sep19.xlsx", sheet = 4)
```

Health Board Codes
```{r}
healthboard_codes <- read_csv("../../data/raw_data/healthboard_codes.csv") %>%
# Clean Names
  clean_names() %>% 
# Convert healthboard area names to match hospitalstays_by_healthboard
  mutate(hb_name =
           str_replace_all(hb_name, " and ", " & ")) %>% 
# Select Relevant Variables
  select(hb_name, hb) %>% 
# Make Readable
  rename("healthboard_area" = "hb")
```

Incidence by Health Board
```{r}
incidence_by_healthboard <- read_xlsx("../../data/raw_data/incidence_byhb.xls", sheet = 3) %>%
  clean_names() %>% 
  rename("sex" = "sex_desc")
```

----
Wrangling
----

# Discharges By Health Board

 -  Add approx year to match mortality
```{r}
# Remove "/[0-9][0-9]" from discharges by hb
discharges_by_healthboard <- discharges_by_healthboard %>%
  mutate(approx_year = as.numeric(str_sub(financial_year, start = 1, end = -4))) %>% 
  select(-financial_year)
```

 - Remove admission type groupings to get the correct number of observations
```{r}
discharges_by_healthboard <- discharges_by_healthboard %>% 
  group_by(approx_year, healthboard_area, age_group, sex, diagnosis) %>% 
  summarise(number_of_discharges = mean(number_of_discharges, na.rm = TRUE),
            crude_rate = mean(crude_rate, na.rm = TRUE),
            easr = mean(easr, na.rm = TRUE)
            )
```

# Hospital Stays by Health Board

 - Filter to only relevant diagnoses, allow missing capital letters
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>%  
  filter(str_detect(lookup, 
      "[Cc]erebrovascular [Dd]isease|[Ss]troke|[Ss]ubarachnoid [Hh]aemorrhage|TIAs|[Tt]ransient [Ii]schemic [Aa]ttack"))
```

 - Seperate lookup column into financial_year, healthboard_area_named, sex, age_group, diagnosis
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>% 
# financial_year
  separate(col = lookup, into = c("financial_year", "lookup_remain1"), sep = 7, remove = FALSE) %>% 
# diagnosis
  separate(col = lookup_remain1, into = c("lookup_remain2", "diagnosis"), sep = " -") %>%
# sex
  mutate(sex = case_when(
    str_detect(lookup, "Male") ~ "Male",
    str_detect(lookup, "Female") ~ "Female",
    str_detect(lookup, "Both Sexes") ~ "All"
  )) %>%
# healthboard_area and age_group
  separate(col = lookup_remain2, into = c("healthboard_area_named", "age_group"), sep = "Male|Female|Both Sexes")  %>% 
# remove original lookup variable from cleaned data
  select(-c(lookup, discharge_finyr, hbresname, age_grp, diag))
```

 -  Add approx year to match mortality
```{r}
# Remove "/[0-9][0-9]" from discharges by hb
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>%
  mutate(approx_year = as.numeric(str_sub(financial_year, start = 1, end = -4))) %>% 
  select(-financial_year)
```

 - Mutate age_groups to match discharges and mortality
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>% 
  mutate(age_group =
    case_when(
      age_group  == "00-14 " ~ "0-44 years",
      age_group  == "15-24 " ~ "0-44 years",
      age_group  == "25-34 " ~ "0-44 years",
      age_group  == "35-44 " ~ "0-44 years",
      age_group  == "45-54 " ~ "45-64 years",
      age_group  == "55-64 " ~ "45-64 years",
      age_group  == "65-74 " ~ "65-74 years",
      age_group  == "65+ " ~ "65-74 years",
      age_group  == "75+ " ~ "75plus years",
      age_group  == "75-84 " ~ "75plus years",
      age_group  == "85+ " ~ "75plus years",
      age_group  == "All ages " ~ "All"
    )
  )
```

 - Mutate diagnosis to match discharges and mortality
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>% 
  mutate(diagnosis = if_else(
    diagnosis == "Cerebrovascular diseases", "Cerebrovascular Disease", diagnosis
  ))

```

 - Remove errant "p" from healthboard areas
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>% 
  mutate(healthboard_area_named = str_remove_all(string = healthboard_area_named, pattern = "^p"))
```

 - Group to get correct number of observations
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard  %>% 
  rename("hospitalstays" = "stays", "crude_rate_stays" = "rate") %>% 
  mutate(hospitalstays = as.numeric(hospitalstays),
         crude_rate_stays = as.numeric(crude_rate_stays)) %>% 
  group_by(approx_year, healthboard_area_named, age_group, sex, diagnosis) %>%
  summarise(hospitalstays = mean(hospitalstays, na.rm = TRUE),
            crude_rate_stays = mean(crude_rate_stays, na.rm = TRUE)
            )
```

# Mortality By Health Board
this code has no effect on current data, but ensures joins will work if new variables and observations are introduced into future versions of this dataset.
```{r}
mortality_by_healthboard <- mortality_by_healthboard %>% 
  group_by(year, healthboard_area, age_group, sex, diagnosis) %>%
  summarise(number_of_deaths = mean(number_of_deaths),
            crude_rate = mean(crude_rate, na.rm = TRUE),
            easr = mean(easr, na.rm = TRUE)
            )
```

# Incidence by Healthboard
```{r}
incidence_by_healthboard <- incidence_by_healthboard %>% 
  filter(sex != "BOTH SEXES") %>% 
  mutate(sex = case_when(
    sex == "MALE" ~ "Male",
    sex == "FEMALE" ~ "Female"
    ),
  age = case_when(
      age  == "0-44" ~ "0-44 years",
      age  == "45-64" ~ "45-64 years",
      age  == "65-74" ~ "65-74 years",
      age  == "75+" ~ "75plus years",
      age  == "ALL AGES" ~ "All",
      age == "<75" ~ "under75 years"
    )
  )
```


----
Joins
----

Health Board Codes and Hospital Stays
```{r}
hospitalstays_by_healthboard <- hospitalstays_by_healthboard %>% 
  left_join(healthboard_codes, 
            by = c("healthboard_area_named" = "hb_name"))
```

Discharges and Mortality
```{r}
strokes_byhb <- discharges_by_healthboard %>% 
  full_join(mortality_by_healthboard, 
            by = c("approx_year" = "year", 
                   "sex" = "sex",
                   "healthboard_area" = "healthboard_area",
                   "age_group" = "age_group",
                   "diagnosis" = "diagnosis"
                   ),
            suffix = c("_discharge", "_mortality")
  )
```

Discharges and Mortality to Hospital stays
```{r}
strokes_byhb <- strokes_byhb %>% 
  full_join(hospitalstays_by_healthboard,
            by = c("approx_year", "age_group", "sex", "healthboard_area", "diagnosis"))
```

Healthboard names to observations without hospital stays data
```{r}
strokes_byhb <- strokes_byhb %>% 
  left_join(healthboard_codes) %>% 
  mutate(healthboard_area_named = if_else(
    is.na(healthboard_area_named),
    hb_name,
    healthboard_area_named
  )) %>% 
# Fix values for readabiliy and matching
  select(-c(hb_name)) %>% 
  mutate(healthboard_area_named = if_else(
    str_detect(healthboard_area_named, "NHS"),
    str_sub(healthboard_area_named, start = 5, end = -1),
    healthboard_area_named
    )
  ) %>% 
  mutate(healthboard_area_named = if_else(
    healthboard_area_named == "cotland",
    "Scotland",
    healthboard_area_named
  ))
```

Ensuring groupings have worked.
```{r}
strokes_byhb <- strokes_byhb %>% 
group_by(approx_year, healthboard_area_named, age_group, sex, diagnosis) %>%
  summarise(
            number_of_discharges = mean(number_of_discharges, na.rm = TRUE),
            crude_rate_discharge = mean(crude_rate_discharge, na.rm = TRUE),
            easr_discharge = mean(easr_discharge, na.rm = TRUE),
            number_of_deaths = mean(number_of_deaths, na.rm = TRUE),
            crude_rate_mortality = mean(crude_rate_mortality, na.rm = TRUE),
            easr_mortality = mean(easr_mortality, na.rm = TRUE),
            hospitalstays = mean(hospitalstays, na.rm = TRUE),
            crude_rate_stays = mean(crude_rate_stays, na.rm = TRUE)
            ) %>% 
  filter(age_group != "All" & age_group != "under75 years") %>% 
  filter(healthboard_area_named != "Scotland"
         & healthboard_area_named != "NA"
         & healthboard_area_named != "Other") %>% 
  filter(sex != "All")
```

----
Write Clean Data
----
```{r}
write_csv(strokes_byhb, "../../data/clean_data/strokes_byhb.csv")

write_csv(incidence_by_healthboard, "../../data/clean_data/incidence_byhb.csv")
```

