----
Libraries
----

```{r}
library(tidyverse)
```

---- 
Data
----
```{r}
council_strokesbyca <- read_csv("../../data/clean_data/strokes_byca.csv",
                                col_types = cols(hospitalstays = col_double(),
                                            crude_rate_stays = col_double()))
healthboard_strokesbyhb <- read_csv("../../data/clean_data/strokes_byhb.csv",
                                col_types = cols(hospitalstays = col_double(),
                                            crude_rate_stays = col_double()))

incidence_by_hb <- read_csv("../../data/clean_data/incidence_byhb.csv")
```

----
Are there any differences in stroke metrics between different demographics (e.g. sex, age group)?
----

Demographics available:
* gender
* age
* healthboard

Outcome variables:
* Diagnosis
* rate of hospital stays
* easr mortalities
* (easr rate discharges - will this tell us anything?)
* Incidence

# Gender

## Stays
```{r}
council_strokesbyca %>% 
  group_by(sex) %>% 
  summarise(hospital_stays = mean(crude_rate_stays, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = sex, y = hospital_stays) +
  geom_col()
```
Men have slightly more hospital stays than women and all genders.

## Mortality
```{r}
council_strokesbyca %>% 
  group_by(sex) %>% 
  summarise(mortality = mean(crude_rate_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = sex, y = mortality) +
  geom_col()
```

```{r}
council_strokesbyca %>% 
  group_by(sex) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = sex, y = mortality) +
  geom_col()
```
Looking at crude rates, women and all genders are recorded with significantly higher mortality rates than males. Though this difference is 10 people per 100,000 of the population or a difference of 0.1%. 
With standardised rates, we see very little difference.

Considering the fact that men are admitted to hospital more often than those with female or all genders and that there is a slightly lower rate of mortality for men, we may want to investigate the impact of hospital admissions on stroke outcomes and test whether the difference between hospital stay rates by gender is significant. With this information, we may conclude that targetting female and all genders with information about when to seek medical help and ensuring that stroke symptoms in women are sufficiently understood.

## Diagnosis
```{r}
council_strokesbyca %>% 
  group_by(sex, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = diagnosis, y = mortality, group = sex, fill = sex) +
  geom_col(position = "fill")
```
Women die from SH proportionally more than men. If we wish to improve mortality outcomes for women we might want to look at treatment and knowledge of SH.

## Incidence

```{r}
incidence_by_hb %>% 
  group_by(sex) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = sex, y = incidence) +
  geom_col()
```
Men have higher rates of incidence than women and other genders, but lower rates of mortality.

## Gender Incidence over time
```{r}
incidence_by_hb %>% 
  group_by(sex, year) %>% 
  summarise(incidence = median(easr, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = year, y = incidence, group = sex, col = sex) +
  geom_line() +
  labs(
    title = "Median incidents by Gender 2009-2018"
  )
```
```{r}
incidence_by_hb %>% 
  group_by(sex, year) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = year, y = incidence, group = sex, col = sex) +
  geom_line() +
  labs(
    title = "Incidents by Gender 2009-2018"
  )
```

With mean: All groups trending in decrease, with female incidence rates being the most volatile. Female incidents rising since 2016, all genders slightly increasing since 2016.

With median: Most significant decrease in 2010, then no significant change. Fluctuation 2010-2014, stasis or increase 2014-2018.

Does this mean we're seeing significant reductions in some demogreaphics which are disproportionally affecting the mean values.

## Gender mortality over time
```{r}
council_strokesbyca %>% 
  group_by(sex, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = sex, col = sex) +
  geom_line() +
  labs(
    title = "Mortality by Gender 2009-2018"
  )
```
Mortalities for male have outstripped females in 2013, 2016 and 2018. All mortality rates declining - yay! Mortality rates for the most recent figures showed male with highest mortalities

# Age

## Stays
```{r}
council_strokesbyca %>% 
  group_by(age_group) %>% 
  summarise(hospital_stays = mean(crude_rate_stays, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = age_group, y = hospital_stays) +
  geom_col()
```


## Mortality
```{r}
council_strokesbyca %>% 
  group_by(age_group) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = age_group, y = mortality) +
  geom_col()
```
Both graphs show us that stroke diagnoses are more common with age, affecting the 75 + age group more significantly than any other group.
This is not surprising information, but the mortality rates for age group do suggest that if our aim is to reduce mortalities by stroke diagnoses, we should target inerventions a the 65+ or 75+ age groups.

## Trends in age group diagnoses over time
stays
```{r}
council_strokesbyca %>% 
  group_by(age_group, approx_year) %>% 
  summarise(hospital_stays = mean(crude_rate_stays, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = hospital_stays, group = age_group, col = age_group) +
  geom_line()
```
This graph shows decline or stasis for hospital admission rates for allage groups between 2014 and 2018. The only noticeable change in hospital stays is a reduction in hospital admissions for the over 75s, which may indicate succesful strategies in these years that could be reproduced for other age groups.

mortality
```{r}
council_strokesbyca %>% 
  group_by(age_group, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = age_group, col = age_group) +
  geom_line()
```
No upwards trends in mortality by stroke. Note that mortality has decreased significantly for the over 75 group so we may be able to identify and copy strategies which have been implemented and extend these to 65 - 74 age group whose mortality has decreased less dramatically since 2009. Alternatively, this may reflect an increase in mortalities from other factors or a change in recording of cause of death. 
 
## Diagnosis
Mortality
```{r}
council_strokesbyca %>% 
  group_by(age_group, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = diagnosis, y = mortality, group = age_group, fill = age_group) +
  geom_col(position = "fill")
```


```{r}
council_strokesbyca %>% 
  group_by(age_group, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = age_group, y = mortality, group = diagnosis, fill = diagnosis) +
  geom_col(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
    title = "Proportion of diagnoses by age group"
  )
```
This shows that if we want to decrease mortalities overall we should focus on CD, but that SH acts differently to the other diagnoses and actually decreases in fatality with age.

discharges
```{r}
council_strokesbyca %>% 
  group_by(age_group, diagnosis) %>%
  summarise(discharges = mean(easr_discharge, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = age_group, y = discharges, group = diagnosis, fill = diagnosis) +
  geom_col(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
    title = "Proportion of discharges by diagnoses and age group"
  )
```

Ratio activity to deaths
```{r}
council_strokesbyca %>% 
  group_by(age_group, diagnosis) %>%
  summarise(discharges = mean(easr_discharge, na.rm = TRUE),
            mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = age_group, y = mortality / discharges * 100, 
        group = diagnosis, fill = diagnosis) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
    title = "Percentage of deaths by hospital discharges",
    subtitle = "shown by diagnoses and age group"
  )
```
Explore SH across time
```{r}
council_strokesbyca %>% 
  filter(diagnosis == "Subarachnoid Haemorrhage") %>% 
  group_by(age_group, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = age_group, col = age_group) +
  geom_line()
```

Mortality by age and diagnosis over time
```{r}
council_strokesbyca %>% 
# filter year to remove high values from years prior to 2013 in order to make graph more readable
  filter(age_group != "All" & age_group != "under75 years" & approx_year > 2013) %>% 
  group_by(age_group, approx_year, diagnosis) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = diagnosis, col = diagnosis) +
  geom_line() +
  facet_wrap(~age_group)
```
We see decrease in mortalities for the 75 plus age group with CD and stroke, but not SH. All other age groups have consistent mortality rates for all diagnoses.

## Incidence
```{r}
incidence_by_hb %>% 
  group_by(age) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = age, y = incidence) +
  geom_col()
```

# Gender and Age

## Stays
```{r}
council_strokesbyca %>% 
  group_by(age_group, sex) %>% 
  summarise(stays = mean(crude_rate_stays, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = age_group, y = stays, group = sex, fill = sex) +
  geom_col(position = "dodge")
```

## Mortality
```{r}
council_strokesbyca %>% 
  group_by(age_group, sex) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = age_group, y = mortality, group = sex, fill = sex) +
  geom_col(position = "dodge")
```
Men have slightly higher mortality rates for all age groups under 74.
For the 75+ group, women have the highest mortality.

## Incidence
```{r}
incidence_by_hb %>% 
  group_by(age, sex) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = age, y = incidence, group = sex, fill = sex) +
  geom_col(position = "dodge")
```
Men have higher incidence rates than women and all genders in all age groups. 

## Percent death by incidence by age and gender
```{r}
percent_table <- incidence_by_hb %>% 
  group_by(age, sex) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  rename("age_group" = "age")

healthboard_strokesbyhb %>% 
  group_by(age_group, sex) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  inner_join(percent_table) %>% 
  ggplot() +
  aes(x = age_group, y = mortality / incidence * 100, group = sex, fill = sex) +
  geom_col(position = "dodge") +
  labs(
    title = "Deaths per 100 incidents",
    subtitle = "grouped by age and gender"
  )
```
From 45 years upwards, women have the highest number of deaths per incident. Under 44 males have the highest number of deaths per incident

## Mortality by Diagnosis
```{r}
council_strokesbyca %>% 
  group_by(sex, age_group, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = age_group, y = mortality, group = diagnosis, fill = diagnosis) +
  geom_col(position = "fill") +
  facet_wrap(~sex) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
We start to see that younger women with SH, especially the 45-64 age group, are dying more often from SH than male and all genders, where men in this age group are more likely to die from stroke.

# Healthboard

## Stays
```{r}
healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named) %>% 
  summarise(hospital_stays = mean(crude_rate_stays, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = reorder(healthboard_area_named, -hospital_stays), y = hospital_stays) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
This shows little variation in hospital admission by health board. The highes admissions being Orkney makes me want to consider access to primary care facilities, while we may wish to see whether any strategies have been used in Grampian and not elsewhere. Unfortunately this data is not normalised for population demographics so we could be looking at differences in demographics rather than stroke experiences.

## Mortality
```{r}
healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = reorder(healthboard_area_named, -mortality), y = mortality) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Differences between health boards of 12.5 deaths per 100,000 population (adjusted). Though this is a small proportion, it does tally with other health inequalities and may encourage some action to target health in Lanarkshire, Highlands and Glasgow.

## Incidence
```{r}
incidence_by_hb %>% 
  group_by(health_board) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  ggplot(aes(x = reorder(health_board, -incidence), y = incidence)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Table of values
```{r}
healthboard_table <- healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named) %>% 
  summarise(hospital_stays = mean(crude_rate_stays, na.rm = TRUE),
            mortality = mean(easr_mortality, na.rm = TRUE)
  ) 

healthboard_table <- incidence_by_hb %>% 
  group_by(health_board) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>%
  mutate(health_board = if_else(
    str_detect(health_board, "NHS"),
    str_sub(health_board, start = 5, end = -1),
    health_board
    )
  ) %>% 
  full_join(healthboard_table, by = c("health_board" = "healthboard_area_named"))
```

incidence and hospital stays
```{r}
healthboard_table %>% 
  ggplot(aes(x = incidence, y = hospital_stays)) +
  geom_point() +
  geom_text(aes(label = health_board), hjust = -.2) +
  labs(
    title = "Comparing incidence rates and hospital visits by Health Board"
  )
```
Island health boards have more hospital stays per incident of stroke than land based health board, but hospital stays are not normalised.

incidence mortality
```{r}
healthboard_table %>% 
  ggplot(aes(x = incidence, y = mortality)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_text(aes(label = health_board), hjust = -.2) +
  labs(
    title = "Mortalities per incident by Health Board"
  )
```
Higher mortalities per incidence in Lanarkshire, Highlands, Grampian and Western Isles.
Lower mortalities per incidence in the Borders, Tayside, Lothian and Fife.

## Mortalities over time by health board

```{r}
# group = health_board, x = year, y = mortality
healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = healthboard_area_named, col = healthboard_area_named) +
  geom_line() +
  geom_point()
  
```
Outliers in 2011, 2013, 2015 and 2018.

Grouping high and low mortality boards for all values, Mean.
```{r}
high_mortality_healthboard <- healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_max(mortality, n = 9) %>% 
  pull(healthboard_area_named)

low_mortality_healthboard <- healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_min(mortality, n = 8) %>% 
  pull(healthboard_area_named)
```

```{r}
healthboard_strokesbyhb %>% 
  filter(healthboard_area_named %in% high_mortality_healthboard |
           healthboard_area_named %in% low_mortality_healthboard) %>% 
  mutate(high_low = if_else(
    healthboard_area_named %in% high_mortality_healthboard, "high", "low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = high_low, col = high_low) +
  geom_line()
```

## Proportion of mortalities by incidents by health board over time
```{r}
percent_table2 <- incidence_by_hb %>% 
  group_by(health_board, year) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  rename("healthboard_area_named" = "health_board",
         "approx_year" = "year") %>% 
  mutate(healthboard_area_named = if_else(
    str_detect(healthboard_area_named, "NHS"),
    str_sub(healthboard_area_named, start = 5, end = -1),
    healthboard_area_named
    )
  )

healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  inner_join(percent_table2) %>% 
  mutate(percent_deathbyincident = mortality / incidence * 100) %>% 
  ggplot() +
  aes(x = approx_year, y = percent_deathbyincident, group = healthboard_area_named, col = healthboard_area_named) +
  geom_line() +
  geom_point()
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Percent deaths over time grouped by high and low mortality boards
```{r}
percent_table2 <- incidence_by_hb %>% 
  rename("healthboard_area_named" = "health_board",
         "approx_year" = "year") %>% 
  mutate(healthboard_area_named = if_else(
    str_detect(healthboard_area_named, "NHS"),
    str_sub(healthboard_area_named, start = 5, end = -1),
    healthboard_area_named
    )
  ) %>% 
  mutate(high_low = if_else(
    healthboard_area_named %in% high_mortality_healthboard, "high", "low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE))


healthboard_strokesbyhb %>% 
  mutate(high_low = if_else(
    healthboard_area_named %in% high_mortality_healthboard, "high", "low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  inner_join(percent_table2) %>% 
  group_by(high_low, approx_year) %>% 
  mutate(percent_deathbyincident = mortality / incidence * 100) %>% 
  ggplot() +
  aes(x = approx_year, y = percent_deathbyincident, group = high_low, col = high_low) +
  geom_line() +
  geom_point()
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Diagnosis
```{r}
healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = healthboard_area_named, y = mortality, group = diagnosis, fill = diagnosis) +
  geom_col(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Higher mortality from stroke in Shetland vs normal distribution of deaths per diagnosis.

# Council

## Stays
```{r}
council_strokesbyca %>% 
  group_by(council_area_named) %>% 
  summarise(hospital_stays = mean(crude_rate_stays, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = reorder(council_area_named, -hospital_stays), y = hospital_stays) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Seeing differences of up to 100 in crude_rate for hospital stays.

## Mortality

```{r}
council_strokesbyca %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = reorder(council_area_named, -mortality), y = mortality) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

These adjusted mortality rates show that there is a difference in stroke mortality between different council areas. We will want to investigate whether East Renfrewshire and East Dunbartonshire have helpful strategies in place and may wish to target interventions at Inverclyde and Renfrewshire.

## Council mortality rates over time for highest and lowest mortality rate councils

10 highest and lowest all time mortality rates, mean
```{r}
high_mortality_council <- council_strokesbyca %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_max(mortality, n = 10) %>% 
  pull(council_area_named)

low_mortality_council <- council_strokesbyca %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_min(mortality, n = 10) %>% 
  pull(council_area_named)
```

```{r}
council_strokesbyca %>% 
  filter(council_area_named %in% high_mortality_council |
           council_area_named %in% low_mortality_council) %>% 
  mutate(high_low = if_else(
    council_area_named %in% high_mortality_council, "high", "low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = high_low, col = high_low) +
  geom_line()
```

Greatest divergences in 2010 and 2015, which are the years after the two government strategies, but these distinctions are less clear with 5 highest and lowest mortality councils.

### 10 highest and lowest mortality councils in 2018, Mean 
```{r}
high_mortality_council_2018 <- council_strokesbyca %>% 
  filter(approx_year == 2018) %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_max(mortality, n = 10) %>% 
  pull(council_area_named)

low_mortality_council_2018 <- council_strokesbyca %>%  
  filter(approx_year == 2018) %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_min(mortality, n = 10) %>% 
  pull(council_area_named)
```



```{r}
council_strokesbyca %>% 
  filter(council_area_named %in% high_mortality_council_2018 |
           council_area_named %in% low_mortality_council_2018) %>% 
  mutate(high_low = if_else(
    council_area_named %in% high_mortality_council_2018, "high", "low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = high_low, col = high_low) +
  geom_line() +
  labs(
    title = "Adjusted mortality rates by council",
    subtitle = "Mortality rates for the 10 highest and lowest adjusted mortality rates from stroke in 2018 using mean"
  )
```

The 10 councils with the highest recent mortality rates showed increase in adjusted mortality in 2013, 2015 and 2018 with an overall downward trend.

The 10 councils with the lowest recent mortality rates show much more stable decline excluding the year 2013. Could this be explaiend by more consistency in healthcare approaches or even a political or financial shift in these areas.

Are these explainable?

Do they correlate with years with big outliers for individual cas?
```{r}
council_strokesbyca %>% 
  group_by(council_area_named, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = council_area_named, col = council_area_named) +
  geom_line()
```
Outliers in 2010, 2011, 2013, 2015, 2016 and 2018.

### 10 highest and lowest all time mortality rates, Median
```{r}
high_mortality_council_median <- council_strokesbyca %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = median(easr_mortality, na.rm = TRUE)) %>% 
  slice_max(mortality, n = 10) %>% 
  pull(council_area_named)

low_mortality_council_median <- council_strokesbyca %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = median(easr_mortality, na.rm = TRUE)) %>% 
  slice_min(mortality, n = 10) %>% 
  pull(council_area_named)
```
```{r}
council_strokesbyca %>% 
  filter(council_area_named %in% high_mortality_council_median |
           council_area_named %in% low_mortality_council_median) %>% 
  mutate(high_low = if_else(
    council_area_named %in% high_mortality_council_median, "high", "low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(mortality = median(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = high_low, col = high_low) +
  geom_line() +
  geom_point()
```

Downward trend in mortalities for both groupings, with all time highest mortality councils decreasing more significantly than lowest.
Seeing greatest divergence in 2011 and 2016.
Looking at small differences here of around 5 deaths per 100,000.
