```{r}
library(tidyverse)
library(janitor)
library(stringr)
library(readxl)
```

# Activity by council

```{r}
activity_by_council <- read_csv("raw_data/stroke_activitybyca.csv") %>% 
  select(-contains("QF")) %>% 
  clean_names() %>% 
  rename(council_area = ca) %>% 
  select(-admission_type)
```
```{r}
activity_by_council %>% 
  summarise(across(.fns = ~sum(is.na(.))))
```


How many diagnoses?
```{r}
activity_by_council %>% 
  count(diagnosis)
```

Correlation diagnosis and admission type?

```{r}

```

Or just one obs per group of characteristics.

```{r}

```

Explore EASR by available variables
```{r}
activity_by_council %>% 
  group_by(financial_year) %>% 
  summarise(avg = mean(easr)) %>% 
  ggplot(aes(x = financial_year, y = avg)) +
  geom_col()
```
Increase in EASR per year, peaking in 16/17 - 17/18

```{r}
activity_by_council %>% 
  group_by(council_area) %>% 
  summarise(avg = mean(easr)) %>% 
  ggplot(aes(x = council_area, y = avg)) +
  geom_col()
```
EASR varies by council area

```{r}
activity_by_council %>% 
  group_by(age_group) %>% 
  summarise(avg = mean(easr)) %>% 
  ggplot(aes(x = age_group, y = avg)) +
  geom_col()
```
EASR higher with age. 

```{r}
activity_by_council %>% 
  group_by(diagnosis) %>% 
  summarise(avg = mean(easr)) %>% 
  ggplot(aes(x = diagnosis, y = avg)) +
  geom_col()
```
Higher EASR for cerebovascular disease and stroke.

# Mortality by council

```{r}
mortality_by_council <- read_csv("raw_data/stroke_mortalitybyca.csv") %>% 
  select(-contains("QF")) %>% 
  clean_names() %>% 
  rename(council_area = ca)
```

Links by ca, age_group(?), year is calendar not financial, sex, diagnosis.

Useful info is number of deaths. Need to get number of deaths for each group, council area, sex, diagnosis and approx. year onto activity?

```{r}
mortality_by_council %>% 
  distinct(year)

activity_by_council %>% 
  distinct(financial_year)
```
Mortality: 2009 - 2018

Activity: 2009/10 - 2018/19

```{r}
# Remove "/[0-9][0-9]" from activity by ca
activity_by_council <- activity_by_council %>%
  mutate(approx_year = as.numeric(str_sub(financial_year, start = 1, end = -4)))
```

```{r}
activity_and_mortality <- activity_by_council %>% 
  full_join(mortality_by_council, 
            by = c("approx_year" = "year", 
                   "sex" = "sex",
                   "council_area" = "council_area",
                   "age_group" = "age_group",
                   "diagnosis" = "diagnosis"
                   ))
```


Find a measure of success? Very dodgy with the mixed years, using number of deaths and number of discharges to estimate number of admissions
```{r}
activity_and_mortality %>% 
  mutate(est_percent_death = number_of_deaths / (number_of_deaths + number_of_discharges))
```


# Diagnosis by CA

(https://www.isdscotland.org/Health-Topics/Hospital-Care/Diagnoses/)
```{r}
diagnosis_by_ca <- read_xlsx("raw_data/Diagnosis-by-Council-Area-of-Residence-Sep19.xlsx", sheet = 4)
# number of hospital stays and rates by 100,00 population for diagnoses by CA, age and gender.
```

Try selecting only columns with Cerebrovascular Disease, Stroke, Subarachnoid Haemorrhage, TIAs and related syndromes.
Use standard language from NHS forms, but allow for missing capitals and include Transient Ischemic Attach as well as TIAs

```{r}
diagnosis_by_ca <- diagnosis_by_ca %>%  
  filter(str_detect(lookup, 
      "[Cc]erebrovascular [Dd]isease|[Ss]troke|[Ss]ubarachnoid [Hh]aemorrhage|TIAs|[Tt]ransient [Ii]schemic [Aa]ttack"))
```


Seperate lookup column into financial_year, council_area_named, sex, age_group, diagnosis

```{r}
joinable_diagnosis_by_ca  <- diagnosis_by_ca %>% 
# financial_year
  separate(col = lookup, into = c("financial_year", "lookup_remain1"), sep = 7, remove = FALSE) %>% 
# diagnosis
  separate(col = lookup_remain1, into = c("lookup_remain2", "diagnosis"), sep = " -") %>%
# sex
  mutate(sex = case_when(
    str_detect(lookup, "Male") ~ "Male",
    str_detect(lookup, "Female") ~ "Female",
    str_detect(lookup, "Both Sexes") ~ "All"
  )) %>%
# council_area and age_group
  separate(col = lookup_remain2, into = c("council_area_named", "age_group"), sep = "Male|Female|Both Sexes")
```


Mutate age group to match activity and mortality data
* This stage loses granularity of data *
```{r}
joinable_diagnosis_by_ca %>% distinct(age_group) %>% pull()
activity_and_mortality %>% distinct(age_group) %>% pull()
```
* assumption that 65+ variable relates to those over 65 and under 75 *

```{r}
joinable_diagnosis_by_ca <- joinable_diagnosis_by_ca %>% 
  mutate(age_group =
    case_when(
      age_group  == "00-14 " ~ "0-44 years",
      age_group  == "15-24 " ~ "0-44 years",
      age_group  == "25-34 " ~ "0-44 years",
      age_group  == "35-44 " ~ "0-44 years",
      age_group  == "45-54 " ~ "45-64 years",
      age_group  == "55-64 " ~ "45-64 years",
      age_group  == "65-74 " ~ "65-74 years",
      age_group  == "65+ " ~ "65-74 years",
      age_group  == "75+ " ~ "75plus years",
      age_group  == "75-84 " ~ "75plus years",
      age_group  == "85+ " ~ "75plus years",
      age_group  == "All ages " ~ "All"
    )
  )
```

Fix diagnosis column

```{r}
joinable_diagnosis_by_ca <- joinable_diagnosis_by_ca %>% 
  mutate(diagnosis = if_else(
    diagnosis == "Cerebrovascular diseases", "Cerebrovascular Disease", diagnosis
  ))

```


Check council_area_named - identifies p error
```{r}
joinable_diagnosis_by_ca %>% 
  distinct(council_area_named)
```

Remove p from start of string
```{r}
joinable_diagnosis_by_ca <- joinable_diagnosis_by_ca %>% 
  mutate(council_area_named = str_remove_all(string = council_area_named, pattern = "^p")) %>% 
  select(-lookup)

# check this returns 34 values
joinable_diagnosis_by_ca %>% 
  distinct(council_area_named)
```


Try to check this is working - can't figure out across
```{r}
# joinable_diagnosis_by_ca %>% 
#   summarise(across(.fns = distinct(.x)))
```

# Convert council_area_named to council_area code
(https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/967937c4-8d67-4f39-974f-fd58c4acfda5/download/ca11_ca19.csv)

```{r}
council_codes <- read_csv("raw_data/council_codes.csv") %>%
  clean_names() %>% 
  select(ca, ca_name) %>% 
  mutate(ca_name =
           str_replace_all(ca_name, " and ", " & ")) %>% 
  rename("council_area" = "ca")
  
#  mutate(ca_name = case_when(
#    ca_name == "Argyll and Bute" ~ "Argyll & Bute",
#    ca_name == "Dumfries and Galloway" ~ "Dumfries & Galloway"
#  ))
```

Remove duplicates codes
```{r}
council_codes <- council_codes %>% 
  distinct(council_area, ca_name)
```

```{r}
unique(council_codes$ca_name)
```


Join codes to diagnosis
```{r}
joinable_diagnosis_by_ca <- joinable_diagnosis_by_ca %>% 
  inner_join(council_codes, 
            by = c("council_area_named" = "ca_name")) 
#  filter(is.na(ca)) %>% 
#  distinct(council_area_named)
#  this fails on Other and Scotland, which is expected.


```

# Join diagnosis to activity and mortality

Check variables match
```{r}
# year
joinable_diagnosis_by_ca %>% distinct(financial_year)
activity_and_mortality %>% distinct(financial_year)
# council area
joinable_diagnosis_by_ca %>% distinct(council_area)
activity_and_mortality %>% distinct(council_area)
# age group
joinable_diagnosis_by_ca %>% distinct(age_group)
activity_and_mortality %>% distinct(age_group)
# sex
joinable_diagnosis_by_ca %>% distinct(sex)
activity_and_mortality %>% distinct(sex)
# diagnosis
joinable_diagnosis_by_ca %>% distinct(diagnosis)
activity_and_mortality %>% distinct(diagnosis)
```


```{r}
diagnosis_activity_and_mortality <- activity_and_mortality %>% 
  inner_join(joinable_diagnosis_by_ca,
            by = c("financial_year", "age_group", "sex", "council_area", "diagnosis"))
```

```{r}
diagnosis_activity_and_mortality %>% 
  filter(is.na(approx_year))

activity_by_council %>% 
  distinct(approx_year)
```

```{r}
# all data
expand_grid(1:10, 1:6, 1:3, 1:37, 1:4)
# expect final data to contain 26,640 rows.

# activity
expand_grid(1:10, 1:6, 1:3, 1:30, 1:4)
# expect 21,600 rows. actual = 95040

# mortality
expand_grid(1:10, 1:6, 1:3, 1:33, c("a", "b", "c"))
# expect 17,820. actual = 17820

# diagnosis
expand_grid(1:5, c("a", "b", "c", "d", "e"), 1:3, 1:37, 1:2)
# expect 5,550. actual = 9,785

diagnosis_by_ca %>% 
  distinct(lookup)
```

