```{r}
library(tidyverse)
library(janitor)
library(stringr)
```

# Activity by council

```{r}
activity_by_council <- read_csv("stroke_activitybyca.csv") %>% 
  select(-contains("QF"))
```

How many admission types?
```{r}
activity_by_council %>% 
  count(AdmissionType)
```

How many diagnoses?
```{r}
activity_by_council %>% 
  count(Diagnosis)
```

correlation diagnosis and admission type
```{r}

```

Explore EASR by available variables
```{r}
activity_by_council %>% 
  group_by(FinancialYear) %>% 
  summarise(avg = mean(EASR)) %>% 
  ggplot(aes(x = FinancialYear, y = avg)) +
  geom_col()
```
Increase in EASR per year, peaking in 16/17 - 17/18

```{r}
activity_by_council %>% 
  group_by(CA) %>% 
  summarise(avg = mean(EASR)) %>% 
  ggplot(aes(x = CA, y = avg)) +
  geom_col()
```
EASR varies by council area

```{r}
activity_by_council %>% 
  group_by(AgeGroup) %>% 
  summarise(avg = mean(EASR)) %>% 
  ggplot(aes(x = AgeGroup, y = avg)) +
  geom_col()
```
EASR higher with age. 

```{r}
activity_by_council %>% 
  group_by(Diagnosis) %>% 
  summarise(avg = mean(EASR)) %>% 
  ggplot(aes(x = Diagnosis, y = avg)) +
  geom_col()
```
Higher EASR for cerebovascular disease and stroke.

# Mortality by council

```{r}
mortality_by_council <- read_csv("stroke_mortalitybyca.csv") %>% 
  select(-contains("QF")) %>% 
  clean_names()
```

Links by ca, age_group(?), year is calendar not financial, sex, diagnosis.

Useful info is number of deaths. Need to get number of deaths for each group, council area, sex, diagnosis and approx. year onto activity?

```{r}
mortality_by_council %>% 
  distinct(year)

activity_by_council %>% 
  distinct(FinancialYear)
```
Mortality: 2009 - 2018

Activity: 2009/10 - 2018/19

```{r}
# Remove "/[0-9][0-9]" from activity by ca
activity_by_council <- activity_by_council %>%
  mutate(join_year = as.numeric(str_sub(FinancialYear, start = 1, end = -4)))
```

```{r}
activity_and_mortality <- activity_by_council %>% 
  full_join(mortality_by_council, 
            by = c("join_year" = "year", 
                   "Sex" = "sex",
                   "CA" = "ca",
                   "AgeGroup" = "age_group",
                   "Diagnosis" = "diagnosis"
                   ))
```


Find a measure of success? Very dodgy with the mixed years, using numberof deaths and number of discharges to estimate number of admissions
```{r}
activity_and_mortality %>% 
  mutate(percent_death = number_of_deaths / (number_of_deaths + NumberOfDischarges)) %>% 
  select(percent_death)
```

