---
output:
  revealjs::revealjs_presentation:
    css: style.css
---

```{r setup, include = FALSE}
# Knitting
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width = '55%')

# Libraries
library(tidyverse)
library(ggthemes)

# Numeric Values
options(digits = 2) 

# Data 
council_strokesbyca <- read_csv("../../../data/clean_data/strokes_byca.csv",
                                col_types = cols(hospitalstays = col_double(),
                                            crude_rate_stays = col_double()))
healthboard_strokesbyhb <- read_csv("../../../data/clean_data/strokes_byhb.csv",
                                col_types = cols(hospitalstays = col_double(),
                                            crude_rate_stays = col_double()))

council_strokesbyca <- read_csv("../../../data/clean_data/strokes_byca.csv",
                                col_types = cols(hospitalstays = col_double(),
                                            crude_rate_stays = col_double()))
healthboard_strokesbyhb <- read_csv("../../../data/clean_data/strokes_byhb.csv",
                                col_types = cols(hospitalstays = col_double(),
                                            crude_rate_stays = col_double()))
incidence_by_hb <- read_csv("../../../data/clean_data/incidence_byhb.csv")

discharges_bysimd <- read_csv("../../../data/clean_data/discharges_bysimd.csv")

mortality_bysimd <- read_csv("../../../data/clean_data/mortality_bysimd2013-2017")
```

## Strokes In Scotland Data Report

**How can Scottish health services most effectively prevent and treat stroke diagnoses?**

This project examines data on stroke diagnosis in Scotland in order to inform decision making on stroke care and intervention by public health bodies.

***

## Background

* Stroke is a disease of the cerebrovascular system, the system which supplies blood to the brain.

* Cerebrovascular disease include stroke, Subarachnoid Haemorhage or Transient Ischemic Attacks (commonly known as mini-strokes).

* The main preventable risk factors for stroke are smoking, high blood pressure, lack of
exercise and a poor diet.

* The treatment and prevention of stroke is a national clinical priority for Scotland

* Scotland has launched major stroke care plans in 2009 and 2014.

***

## Data and Methodology

* This project uses publicly available published by Public Health Scotland

* Available data covers the years 2009-2018

* The data includes information on hospital activity (stays and  discharges), mortality rates and incidence rates.

* Demographic detail varies by dataset and include age, gender, Council Area or Health Board and SIMD (Scottish Index of Multiple Deprivation)

* Data is cleaned, analysed and presented in R using Rstudio and Tidyverse.

***

## Key assumptions

* Financial year can be compared to calendar year without significantly undermining the analysis.

* Ambiguously named and defined crude rates in individual datasets refer to the crude rate of the dataset's subject variable.

* Definition of all genders group varies with dataset, so it has been removed from analysis.

* NA, Other and Scotland are not useful values for measures by Health Board or Council Area and have been removed from analysis.

* Cerebrovascular Disease is not a catch-all for stroke, TIAs and SH and therefore contains useful information on unidentified diagnosis.

***

### How do mortality rates vary with diagnosis and demographics? Age

```{r out.with = '80%'}
council_strokesbyca %>% 
# filter year to remove high values from years prior to 2013 in order to make graph more readable
  filter(approx_year > 2013) %>% 
  filter(diagnosis != "TIAs and related syndromes") %>% 
  mutate(age_group = if_else(age_group == "75plus years", "75+", "under 75")) %>% 
  group_by(age_group, approx_year, diagnosis) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = diagnosis, col = diagnosis) +
  geom_line() +
  facet_wrap(~age_group) +
  labs(
    title = "Mortality by age and diagnosis over time",
    x = "Year",
    y = "Mortality (EASR)",
    col = "Diagnosis"
  ) +
  scale_color_calc()
```

* Mortality rates have fallen for the 75+ group with Cardiovascular Disease and Stroke.
* Mortality rates have remained static for all other age and diagnosis groupings.

***

### How do mortality rates vary with diagnosis and demographics? Gender

```{r}
council_strokesbyca %>% 
  filter(diagnosis != "TIAs and related syndromes") %>% 
  group_by(sex, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = diagnosis, y = mortality, group = sex, fill = sex) +
  geom_col(position = "fill") +
  labs(
    title = "Proportion of mortalities for each diagnosis",
    subtitle = "Separated by gender",
    x = "Diagnosis",
    y = "Mortality (EASR)",
    fill = "Gender"
  ) +
  scale_fill_calc()
```

* Women die from Subarachnoid Haemorhage proportionally more than men.

***

### How do mortality rates vary with diagnosis and demographics? Age and Gender

```{r}
council_strokesbyca %>% 
  group_by(sex, age_group, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = age_group, y = mortality, group = diagnosis, fill = diagnosis) +
  geom_col(position = "fill") +
  facet_wrap(~sex) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
    title = "Proportion of mortalities by age, diagnosis and gender",
    x = "Age Group",
    y = "Mortality (EASR)",
    fill = "Diagnosis"
  )  +
  scale_fill_calc()
```

* Younger women with Subarachnoid Haemorhage are dying more often from Subarachnoid Haemorhage than males.
* Under 65, men have higher proportion of mortality by stroke than women

***

### How do stroke outcomes vary with demographics? Gender

```{r}
incidence_by_hb %>% 
  group_by(sex, year) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = year, y = incidence, group = sex, col = sex) +
  geom_line() +
  geom_point() +
  labs(
    title = "Incidents by Gender 2009-2018",
    x = "Year",
    y = "Incidents (EASR)",
    col = "Gender"
  ) +
  scale_color_calc()
```
  
* Men have higher rates of incidence than women.
* Female incidence rates are the most volatile and have been rising since 2016.
* Although men have higher incidence rates than women, they have similar levels of mortality. i.e. more women die per episode of stroke diagnosis than men.

***

### How do stroke outcomes vary with demographics? Age

```{r}
council_strokesbyca %>%
  group_by(age_group, diagnosis) %>%
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
    aes(x = age_group, y = mortality, group = diagnosis, fill = diagnosis) +
  geom_col(position = "fill") +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(
    title = "Proportion of mortalities by diagnoses and age group",
    x = "Age Group",
    y = "Mortality (EASR)",
    fill = "Diagnosis"
  ) +
  scale_fill_calc()
```

* Stroke discharges increase with age group, but account for consistent mortality proportion.
* CD is the highest killer for and proportion of deaths increases with age.
* SH causes a higher proportion of fatalities in younger age groups.

***

### How do stroke outcomes vary with demographics? Gender and Age

```{r}
percent_table <- incidence_by_hb %>%
  group_by(age, sex) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>% 
  rename("age_group" = "age")

healthboard_strokesbyhb %>%
  group_by(age_group, sex) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  inner_join(percent_table) %>% 
  ggplot() +
  aes(x = age_group, y = mortality / incidence * 100, group = sex, fill = sex) +
  geom_col(position = "dodge") +
  labs(
    title = "Deaths per 100 incidents",
    subtitle = "Grouped by age and gender",
    x = "Age group",
    y = "Rate of Mortality by Incident (%)",
    fill = "Gender"
  ) +
  scale_fill_calc()
```

* Mortality rates: under 75 - men highest, over 75 women highest.
* Men have the highest incidence rates for all age groups.
* Deaths per incident: under 45 - men highest, over 45 - women highest.

***

### How do stroke outcomes vary with demographics? Healthboard

```{r}
healthboard_table <- healthboard_strokesbyhb %>% 
  group_by(healthboard_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) 

healthboard_table <- incidence_by_hb %>% 
  group_by(health_board) %>% 
  summarise(incidence = mean(easr, na.rm = TRUE)) %>%
  mutate(health_board = if_else(
    str_detect(health_board, "NHS"),
    str_sub(health_board, start = 5, end = -1),
    health_board
    )
  ) %>% 
  full_join(healthboard_table, by = c("health_board" = "healthboard_area_named")) %>% 
  arrange(desc(incidence))
  
healthboard_table %>%
  ggplot(aes(x = incidence, y = mortality)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "navy") +
  geom_text(aes(label = health_board), hjust = -.2) +
  labs(
    title = "Mortalities per incident by Health Board",
    x = "Number of Incidents (EASR)",
    y = "Number of Mortalities (EASR)"
  )
```

* Little variation in mortality, up to 100 difference in incidents.
* Higher mortalities per incidence in Lanarkshire, Highlands, Grampian and Western Isles.
* Lower mortalities per incidence in the Borders, Tayside, Lothian and Fife.

***

### How do stroke outcomes vary with demographics? Council

```{r}
high_mortality_council_2018 <- council_strokesbyca %>% 
  filter(approx_year == 2018) %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_max(mortality, n = 10) %>% 
  pull(council_area_named)

low_mortality_council_2018 <- council_strokesbyca %>%  
  filter(approx_year == 2018) %>% 
  group_by(council_area_named) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  slice_min(mortality, n = 10) %>% 
  pull(council_area_named)

council_strokesbyca %>% 
  filter(council_area_named %in% high_mortality_council_2018 |
           council_area_named %in% low_mortality_council_2018) %>% 
  mutate(high_low = if_else(
    council_area_named %in% high_mortality_council_2018, "High", "Low"
  )) %>% 
  group_by(high_low, approx_year) %>% 
  summarise(mortality = mean(easr_mortality, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = approx_year, y = mortality, group = high_low, col = high_low) +
  geom_line() +
  geom_point() +
  labs(
    title = "Adjusted mortality rates by council",
    subtitle = "Mortality rates for the 10 highest and lowest adjusted mortality rates council areas",
    x = "Year",
    y = "Mortality (EASR)",
    col = "Grouping"
  ) +
  scale_color_calc()
```

* Highest adjusted mortality in Inverclyde and Renfrewshire.
* Lowest adjusted mortality rates in East Renfrewshire and East Dunbartonshire.
* The 10 councils with the lowest recent mortality rates show much more stable decline.

***

### How do stroke outcomes vary with demographics? SIMD

```{r}
discharges_bysimd %>% 
  mutate(approx_year = as.character(approx_year)) %>% 
  group_by(deprivation_quintile, approx_year) %>% 
  summarise(avg_discharges = mean(easr_discharge, na.rm = T)) %>% 
  ggplot(aes(x = approx_year, y = avg_discharges, group = deprivation_quintile, col = deprivation_quintile)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Hospital activity by SIMD quintile over time",
    x = "Year",
    y = "Discharges (EASR)",
    col = "SIMD Quintile"
  )
```

<br>

* Hospital activity increases with deprivation.
* Difference between hospital activity rates for deprivation quintiles is increasing over time, particularly for the 2 most deprived quintiles.

***

### Table showing standardised mortality rates by SIMD decile
```{r}
mortality_bysimd %>% 
  group_by(simd_decile) %>% 
  summarise(mortality = mean(smr_mortality)) %>% 
    ggplot(aes(x = simd_decile, y = mortality, fill = "navy")) +
    geom_col() +
    scale_fill_calc() +
  labs(
    title = "Mortality by SIMD Decile",
    x = "Mortality (EASR)",
    y = "SIMD Decile"
  ) +
  theme(legend.position = "none")
```
<br>

* Mortalities increase with deprivation.
* Chance of dying from stroke doubles from the least to the most deprived deciles.

***

### How do stroke outcomes vary with demographics? Age Group and SIMD

```{r}
mortality_bysimd %>% 
  group_by(simd_decile, age_group) %>% 
  filter(age_group != "All") %>% 
  summarise(avg_mortality = mean(smr_mortality, na.rm = T)) %>% 
  ggplot(aes(x = simd_decile, y = avg_mortality, group = age_group, fill = age_group)) +
  geom_col(position = "dodge") +
  labs(
    title = "Adjusted stroke mortality rates by SIMD",
    x = "SIMD Decile",
    y = "Mortality (SMR)",
    fill = "Age Group"
  ) +
  scale_fill_calc()
```

* SIMD has the highest impact on mortality for the under 65 age group. 
* 189.5 people per 100,000 under 65 (adjusted) die of strokes in the most deprived decile, compared to just 42.3 of those in the least deprived decile.

***

### What measures have been successful in the treatment and prevention of stroke?

Direct answer unavailable in the available data, but useful information obtained

**Recommendations for future research:**

* Research interventions that have targetted the 75+ age group, particularly with Stroke and Cardiovascular Disease

* Explore the impact of council policies and politics on health measures.

***

### Which demographics should we target for stroke intervention?

**Future research:**

* Explore data collection to find diagnosis which accounts for excess figures in CD diagnosis.
* Model social factors which contribute to stroke incidence and mortality, such as smoking, high blood pressure, caring duties, mental wellbeing.
* Compare diagnosis rates on presentation and discharge hospital to identify whether misdiagnosis of Subarachnoid Haemorhage is more common in women or younger age groups.
* Conduct research on the understanding and identification of Subarachnoid Haemorhage in women.
* Generate adjusted figures for hospital activity to explore hospital attendance by demographic.

**Recommended interventions:**

* Targetted interventions for women and younger people in deprived areas.
* Publicise stroke symptoms with an emphasis on women and hospital attendance


## Further information on data and methodology available on request
